
# PS 4 Solutions {-}

&nbsp;


## Problem 1 {-}

In this problem use functions from the `apply()` family of functions whenever you think you need `loops`.

&nbsp;

1. Create a function (name it `fun1`) that takes three arguments _n_, _miu_, and _sd_ (where _n_ is the number of randomly generated numbers, _miu_ is the mean value of the distribution that numbers come from, and _sd_ is the standard deviation of that distribution) and does the following:

   * First it generates _n_ random numbers from a normal distribution with mean = _miu_ and standard deviation = _sd_
   * Then it populates an empty matrix as follows: if _n_ is divisible by 2 but not by 3, it fills a matrix by row, having 2 rows and _n/2_ columns; if _n_ is divisible by 3 but not by 2, it fills a matrix by column, having 3 columns and _n/3_ rows; if _n_ is divisible by 2 and by 3, it fills a matrix by row, having 6 rows and _n/6_ columns; else it fills by row and has only one column
   * And returns a vector of mean values of each column in the matrix if none of the randomly generated numbers are negative; it returns a vector of median values of each row in the matrix if some of the randomly generated numbers are negative.
  
  
   Test your function with (_n_ = 50, _miu_ = 0.5, _sd_ = 2)
   
```{r}
set.seed(1)

fun1 <- function(n, miu, sd){
  
  data <- rnorm(n, mean=miu, sd=sd)
  
  if (n%%2==0 && n%%3!=0) {
    
    mat <- matrix(data, byrow=T, nrow=2)
    
  }
  
  else if (n%%3==0 && n%%2!=0) {
    
    mat <- matrix(data, byrow=F, ncol=3)
    
  }
  
  else if (n%%2==0 && n%%3==0) {
    
    mat <- matrix(data, byrow=T, nrow=6)
    
  }
  
  else {
    
    mat <- matrix(data, byrow=T, nrow=1)
    
  }
  
  print("The matrix is:")
  
  print(mat)
  
  if (all(mat>=0)) {
    
    result <- apply(mat, 2, mean)
    
  }
  
  else {
    
    result <- apply(mat, 1, median)
    
  }
  
  print("The funciton result is:")
  
  return (result)
}

fun1(n = 50, miu = 0.5, sd = 2)
```

&nbsp;

2. Create a function (name it `fun2`) that will take one argument _n_, which represents a number of trials in your simulation. Your function should simulate a process of rolling an unfair die with the following probability distribution: "1" - 0.1, "2" - 0.3, "3" - 0.2, "4" - 0.05, "5" - 0.3, "6" - 0.05. It should report the results of _n_ trials by returning the frequency table of the outcomes (for example, how many times you got "1", "2", and so on).

   Test your function with _n_ = 100.

   
"1” - 0.1, “2” - 0.3, “3” - 0.2, “4” - 0.05, “5” - 0.3, “6” - 0.05

```{r}
set.seed(1)

fun2 <- function(n) {
  
  trials <- sample(1:6, n, replace=T, prob=c(0.1, 0.3, 0.2, 0.05, 0.3, 0.05))
  
  return(table(trials))
  
}

fun2(100)
```
   
  &nbsp; 
  
3. Create a function (name it `fun3`) that will take one argument _n_, which represents a number of trials in your simulation. Your function should simulate a process of rolling two fair dice. It should report the results of _n_ trials by returning vectors of outcomes from both dice (name them `Die1` and `Die2`, respectively) and a vector that compares outcomes obtained from `Die1` and `Die2` (name it `comparison`). `comparison` vector should be created as follows: it should be created using `mapply()` function and should be a factor with three levels ("Die1 > Die2", "Die1 < Die2", "Die1 = Die2"). For instance, if the outcome of `Die1` is 5 and the outcome of `Die2` is 4, then it should produce "Die1 > Die2". 

   Test your function with _n_ = 30.
   
```{r}
set.seed(1)

fun3 <- function(n) {
  
  Die1 <- sample(1:6, n, replace=T)
  
  Die2 <- sample(1:6, n, replace=T)
  
  compare <- function(a, b) {
    
    if (a>b) {
      
      return (factor("Die1 > Die2"))
    }
    
    else if (a<b) {
      
      return (factor("Die1 < Die2"))
      
    }
    else {
      
      return (factor("Die1 = Die2"))
      
    }
    
  }
  
  comparison <- mapply(compare, Die1, Die2)
  
  return (data.frame(Die1, Die2, comparison))
  
}

fun3(30)
```

  &nbsp;
  
  
## Problem 2 {-}

```{r}

data <- read.csv("ames_housing.csv", header = T, stringsAsFactors = TRUE)

library(tidyverse)

```

1. John is a new real estate agent in the area. A friend of his, who worked in this area before, told him that the average price for houses sold in the area is \$210,000. John feels a bit skeptical about this claim and wants to test it. Perform an appropriate statistical procedure to test this claim. Use the $\alpha = 0.05$ significance level to make a conclusion. State the null and alternative hypotheses. What is the test statistic? What are the degrees of freedom? What is the p-value? What did you conclude?

\begin{align}
& H_{0}: \textrm{The average price for houses sold in the area equals \$210,000.} \\
& H_{1}: \textrm{The average price for houses sold in the area } \neq \textrm{ \$210,000.}
\end{align}

The test statistics is $t = \frac{\bar{x}-210000}{s/sqrt{n}} \sim t(n-1)$, where $n$ is the sample size:

```{r}
length(data$Sale_Price)
```

The degrees of freedom equals $n-1 = 2924$.

```{r}
t.test(x=data$Sale_Price,
       alternative = "two.sided",
       mu = 210000,
       conf.level = 0.95)
```

Based on the t-test, the $p$-values is $<`r 2.2e-16`$. Thus, we reject the null hypothesis and conclude that "The average house price in the area is different from \$210,000".

&nbsp;

2. John has been in this industry for quite some time. Over the course of past 10 years, he's noticed that a type of roof that houses have can somehow define the fence quality. In other words, he thinks that these two features (`Roof_Style` and `Fence` variables in the dataset) are somehow related. Test his hypothesis based on the sample data at the  $\alpha = 0.01$ significance level. State the null and alternative hypotheses. What is the test statistic? What are the degrees of freedom? What is the p-value? What did you conclude?


\begin{align}
& H_{0}: \textrm{"Roof Style" and "Fence" are independent. The observed difference is due to chance.} \\
& H_{1}: \textrm{"Roof Style" and "Fence" are not independent. The observed difference is not due to chance.}
\end{align}

The test statistics is
$$
\chi^{2} = \sum_{i=1}^{k}\frac{(\textrm{observed}_{i} - \textrm{exptected}_{i})^{2}}{\textrm{expected}_{i}}.
$$

The observed values are cell values in the table:

```{r}
data %>%
  
  select(Roof_Style, Fence) %>%
  
  table() %>%
  
  addmargins()
```

Suppose "Roof\_Style" and "Fence" are independent. The expected value of "Good\_Privacy" fence and "Flat" roof style should be:

$$
118\times\frac{20}{2925} \approx 0.8068.
$$

The degrees of freedom is:
$$
(\textrm{number of rows} - 1)(\textrm{number of columns}-1) = (5-1)\times(5-1) = 16.
$$

```{r warning=FALSE}
chisq.test(data$Roof_Style, data$Fence)
```

The $p$-value is $0.1522$. It is larger than the significance level $0.1$. We fail to reject the null hypothesis. We conclude that the roof style and the fence are independent.

&nbsp;

3. John is specifically interested in two neighborhoods: `North_Ames` and `College_Creek`. He did some research prior his arrival and he thinks that on average houses in the `College_Creek` neighborhood are \$50,000 more expensive than the houses sold in the `North_Ames` neighborhood. Test his hypothesis at the $\alpha = 0.02$ significance level. State the null and alternative hypotheses. What is the test statistic? What are the degrees of freedom? What is the p-value? What did you conclude?

Let $\mu_{\textrm{collegeCreek}}$ be the mean housing price in "College\_Creek" and $\mu_{\textrm{northAmes}}$ be the mean housing price in "North\_Ames". Then the hypotheses are:

\begin{align*}
& H_{0}: \mu_{\textrm{collegeCreek}}-\mu_{\textrm{northAmes}}  =  \textrm{\$50,000}. \\
& H_{1}: \mu_{\textrm{collegeCreek}}-\mu_{\textrm{northAmes}} > \textrm{\$50,000}.
\end{align*}

Perform Levene's test to test if the variances of housing price in "College\_Creek" and that in "North\_Ames" are homogeneous.

```{r message=FALSE}
library(car)
data1 <- data %>% filter(Neighborhood %in% c("College_Creek", "North_Ames"))
leveneTest(Sale_Price ~ Neighborhood, data = data1)
```

$p$-value from Levene's test is pretty small, indicating the variances of the housing price in "College\_Creek" and that in "North\_Ames" are different. 

The  test statistics is 
$$
t = \frac{\bar{x}_{1}-\bar{x}_{2}-50000}{\sqrt{s_{1}^{2}/n_{1}+s^{2}_{2}/n_{2}}} \sim t(n_{1}+n_{2}-2).
$$

The degrees of freedom is:
```{r}
nrow(data1)-2
```


```{r}
collegeCreek <- data1 %>%
  
         filter(Neighborhood == "College_Creek") %>%
          
         select(Sale_Price)

northAmes <- data1 %>%
  
         filter(Neighborhood == "North_Ames") %>%
          
         select(Sale_Price)

t.test(x = collegeCreek,
       
       y = northAmes,
       
       alternative = "greater",
       
       var.equal = FALSE,
       
       mu = 50000,
       
       conf.level = 0.98)
```

The $p$-value from the two-sample t test is $0.033$, which is greater than the significance level $0.02$. We fail to reject $H_{0}$. We conclude that the average housing price in "College\_Creek" is not \$50,000 more expensive than that in "North\_Ames".


&nbsp;

4. John decides to split all homes in the dataset into the following three groups based on the total number of rooms above the ground: `"2-4"`, `"5-8"`, and `"9 or more"`. He believes that 10\% of homes belong to the `"2-4"` category, 85\% belong to the `"5-8"` category, and the remaining 5\% belong to the `"9 or more"` category. Test his hypothesis using the appropriate statistical procedure at the $\alpha = 0.01$ significance level. State the null and alternative hypotheses. What is the test statistic? What are the degrees of freedom? What is the p-value? What did you conclude?

   Before you perform the test, you need to add a new variable to the dataset (name it `Rooms`) using the splitting rule proposed by John. In other words, you need to create a factor variable with 3 levels by converting the `TotRms_AbvGrd` variable. (Hint: you might want to check out the `case_when()` function from `dpyr` package).


```{r}
categorize <- function(x) {
  case_when(x<=4 & x>=2 ~ "2-4",
            x<=8 & x>=5 ~ "5-8",
            x>=9 ~ "9 or more") %>%
  as.factor()
}
data <- data %>%
  mutate(Rooms = categorize(TotRms_AbvGrd))
```

\begin{align*}
& H_{0}: \textrm{The distribution of the number of rooms above the ground is:} \\
& \textrm{"2-4" accounts for 10%, "5-8" accounts for 85% and "9 or more" accounts for 5%.} \\
& H_{1}: \textrm{The distribution of the number of rooms above the ground is NOT: }\\
& \textrm{"2-4" accounts for 10%, "5-8" accounts for 85% and "9 or more" accounts for 5%.}
\end{align*}

The test statistics is
$$
\chi^{2} = \sum_{i=1}^{k}\frac{(\textrm{observed}_{i} - \textrm{exptected}_{i})^{2}}{\textrm{expected}_{i}}.
$$
The degrees of freedom is $3-1 = 2$.

```{r}
data %>% 
  select(Rooms) %>%
  table()
```

```{r}
chisq.test(x=c(228, 2424, 273), p=c(0.1, 0.85, 0.05))
```

The $p$-value is $<`r 2.2e-16`$, smaller than the significance level $0.01$. We reject the null hypothesis and conclude that the distribution of the grades of the number of rooms above ground is not: "2-4" ~ 10\%, "5-8" ~ 85\% and "9 or more" ~ 5\%.



&nbsp;


## Problem 3 {-}


1. Fit the least squares linear regression model. 

```{r}

data2 <- read.table(file = "copier_maintenance.txt", header = T, sep = "")


model <- lm(Minutes ~ Copiers, data = data2)

```

&nbsp;

2. Display the summary of the model obtained in part 1.

```{r}

summary(model)

```

&nbsp;

3. Plot the data and overlay the linear regression function you obtained in part 1.

```{r}

library(tidyverse)

  ggplot(data = data2, aes(x = Copiers, y = Minutes)) + 
  
  geom_point(color = "steelblue") +
  
  labs(title = "Minutes versus Copiers",
       
       x = "Copiers",
       
       y = "Minutes") +
  
  geom_smooth(method = 'lm', se = FALSE)

```

&nbsp;

4. Obtain a point estimate of the mean service time when 7 copiers are serviced. 

```{r}

predict(model, data.frame(Copiers = 7))

```


&nbsp;

5. Obtain a 95% confidence interval for the slope.

```{r}

confint(model, "Copiers", level = 0.95)

```

&nbsp;

6. Obtain a 95% confidence interval for the mean service time when 7 copiers are serviced.

```{r}

predict(model, data.frame(Copiers = 7), interval = "confidence", level = 0.95)

```

&nbsp;

7. Obtain a 95% prediction interval for the mean service time when 7 copiers are serviced.

```{r}

predict(model, data.frame(Copiers = 7), interval = "prediction", level = 0.95)

```